Formal Modelling of Predator Preferences using Molecular Gut-Content Analysis

### 2015-11-18 paper/main.tex
Simon's edits.

\documentclass[smallextended]{svjour3}
\input{TexOpts.tex}
\begin{document}
\title{Formal Modelling of Predator Preferences using Molecular Gut-Content Analysis}
\author{Edward A. Roualdes \and Simon J. Bonner \and Thomas D. Whitney \and James D. Harwood}
\institute{E. A. Roualdes \and S. J. Bonner
        \at Department of Statistics, University of Kentucky, Rm.~311 Multidisciplinary Science Building, 725 Rose Street, Lexington, KY 40536-0082, \\\email{edward.roualdes@uky.edu} \and
        T.D. Whitney 
        \at Department of Entomology, University of Kentucky, Lexington, KY 40546, USA \emph{Present address: Warnell School of Forestry and Natural Resources, University of Georgia, Athens, GA 30602, USA}
        \and
        J.D. Harwood
        \at Department of Entomology, University of Kentucky, Lexington, KY 40546, USA
}

@@

A comprehensive overview by \citet{Lechowicz:1982}, later summarised by \citet{Manly:2002}, details some of the benefits and faults of the most popular indices.  According to these reviews, a majority of the indices give comparable results, save the linear index $L$ by \citet{Strauss:1979}, despite the fact that most of the methods differ by range and linearity of response.  While Lechowicz recommends one index, $E^*$ by \citet{Vanderploeg:1979} as the ``single best'' \citep{Lechowicz:1982}, albeit imperfect, index, Manly et al.\ instead take the approach of excluding the subset of indices which do not ``estimate any biologically meaningful value'' \citep{Manly:2002}.  \citet{Lechowicz:1982} recommends the index $E^*$, an element of the \citet{Manly:2002} suggested indices, because the index value $0$ denotes random feeding, the index has a range restricted to $[-1,1]$ (though $E^*=1$ is nigh impossible), and the index is based on the predator's choice of prey as a function of both the availability of the prey as well as the number of available prey types (assumed known).  The downside to this index is its lack of reasonable statistical properties \citep{Lechowicz:1982}, thus making the computation of standard errors and thus hypothesis testing difficult.  This is, in fact, a common fault amongst most of the indices.

@@

Our model assumes that numbers of each prey species captured and the numbers consumed by an individual predator follow independent Poisson distributions with separate rates that might possibly vary between species and over time.  Further, we are able to estimate the rates of interest even when exact counts of each prey species eaten within any given time period are not observed.  When exact counts are unobserved, we instead rely on the researcher being able to detect prey DNA within the predator's gut \citep{Schmidt:2014,Raso:2014,Madduppa:2014} and make a simple binary conclusion: this predator ate some of that prey species during this time period, or did not.

@@

We assume that samples of both the predators and prey are captured from the study area on $T$ occasions.  Depending on the species involved and the design of the study the predators and prey may be sampled in the same way or using different methods.  In the spider experiment described in Section~\ref{sec:data}, for example, prey are captured using pitfall traps that are dispersed throughout the study area whereas the predators (spiders) were captured by hand.  Predators and prey species are collected and counted at each time period.  We denote the number of predators and the number of prey species caught in each time period $t \in \{1, \ldots, T\}$ by $J_t$ and $I_t$, respectively.  Prey species will be indexed by $s \in \{1, \ldots, S \}$.  Let $X_{jst}$ represent the number of prey species $s$ that predator $j$ ate during period $t$, where $j \in \{1, \ldots, J_t\}$.  Let $Y_{ist}$ represent the number of prey species $s$ found in trap $i$ during period $t$, $i \in \{1, \ldots, I_t\}$.

@@

\begin{enumerate}
\item $c_{st} = 1$
\item $c_{st} = c$
\item $c_{st} = c_s$
\item $c_{st} = c_t$
\item $c_{st} = c_{st}$
\end{enumerate}

\noindent The first hypothesis states that the relative rate of sampling for the predator and the traps is the same for all species on all occasions.  One imagines this is the case if the prey move randomly and the predator simply eats prey which comes within its reach, thus suggesting no selection for a particular prey item.  The second states that predators sample prey proportionally across all time periods.  The third hypothesis states that predators sample different prey species at different rates, but each rate is steady across time.  This implies that the predator expresses preferences for one prey species over another, but is unresponsive to changes due to time.  Conversely, the fourth hypothesis implies that each prey species is sampled similarly within each time period, while the rates across time are allowed to change.  The fifth hypothesis assumes a predator's selection varies by both time and prey species.  This would make sense if environmental and biological variables, such as weather, prey availability, and/or palatability were affecting predators' selection strategies.

@@

Because the five hypotheses are nested, a natural testing order is suggested in Figure~\ref{fig:hier}.  For instance, suppose interest lies in the predator's dietary preferences with respect to prey species $s_1$ and $s_2$ across three time periods.  The hypothesis that the predator has the same preference for the two prey species can be tested with

\begin{align*}
  H_0&: \lambda_t = c_t \gamma_t, \forall t \\  
  H_1&: \lambda_{st} = c_{st}\gamma_{st}.
\end{align*}

The alternative here allows for the predators preferences to vary independently for the two species and across the three time points. Similarly, the hypotheses

\begin{align*}
  H_0&: \lambda_s = c_s \gamma_s, \forall s \\  
  H_1&: \lambda_{st} = c_{st}\gamma_{st}
\end{align*}

would test if the predator's eating varies by prey species, but not by time, against the most parameter rich model $c_{st}$.  The above two tests should be performed before any of the simpler methods are tested.

@@

\noindent Writing all five hypotheses as $\lambda_{st} = c_{st}\gamma_{st}$, we can, in the simplest cases, find analytic solutions for the maximum likelihood estimates (MLEs) of $c_{st}$, $\gamma_{st}$, and by invariance $\lambda_{st}$.  Such solutions exist for the simplest model, $c_{st}=1$, for which only $\gamma_{st}$ need be estimated, and for the second simplest model, $c_{st}=c$, provided that the data are balanced so that $J_t=J$ and $I_t=I$.  Respectively, these solutions are

\begin{equation*}
  \hat{\gamma}_{st} = \frac{X_{\cdot st} + Y_{\cdot st}}{J_t + I_t}, \quad \text{ and } \quad \hat{c} = \frac{I \sum_{s,t} X_{\cdot st}}{J \sum_{s,t} Y_{\cdot st}}, \quad \hat{\gamma}_{st} = \frac{X_{\cdot st} + Y_{\cdot st}}{I \left( \frac{\sum_{st} X_{\cdot st}}{\sum_{st} Y_{\cdot st}} + 1 \right)},
\end{equation*}

\noindent where $X_{\cdot st} = \sum_{j=1}^{J_t}X_{jst}$ and $Y_{\cdot st} = \sum_{i=1}^{I_t} Y_{ist}$.

@@

We denote the binary random variable indicating whether the $j^{th}$ predator did in fact eat at least one individual of prey species $s$ in time period $t$ by $Z_{jst} = 1(X_{jst} > 0)$.  Given the Poisson assumptions above, these variables are independent Bernoulli observations with success probability $p_{st} = P(Z_{jst}=1)= 1-\exp\{-\lambda_{st}\}$.  Despite not observing $X_{jst}$, we can compute maximum likelihood estimates of the parameters $\boldsymbol{\lambda}, \boldsymbol{\gamma}$ through the EM algorithm using the complete data log-likelihood
\[
l_{comp}(\boldsymbol{\lambda}, \boldsymbol{\gamma}) = \log f_{X,Y,Z}(\boldsymbol x, \boldsymbol y, \boldsymbol z|\boldsymbol{\lambda}, \boldsymbol{\gamma}) = \sum_{s,t}^{S,T} \left[ \sum_{j=1}^{J_t} \log f_{X,Z}(x_{jst},z_{jst}|\boldsymbol{\lambda}) + \sum_{i=1}^{I_t}\log f_Y(y_{ist}|\boldsymbol{\gamma}) \right].
\]

% !143, 146
The mass function of $Y_{jst}$ is exactly as in Section~\ref{sec:count} and so we focus on deriving the joint probability mass function of $X_{jst}$ and $Z_{jst}$.  With the distribution of $Z_{jst}$ given above, we can compute $f_{X,Z}(x_{jst},z_{jst}|\boldsymbol{\lambda})$ by noting that $X_{jst}=0$ with probability 1 if $Z_{jst}=0$, and that $[X_{jst}|Z_{jst}=0]$ has a truncated Poisson distribution with mass function
\[
  f_{X|Y,Z,\boldsymbol{\lambda},\boldsymbol{\gamma}}(x_{jst}|z_{jst}) =
  \frac{\exp{\{-\lambda_{st}\}} \lambda_{st}^{x_{jst}}}{(1 - \exp{\{-\lambda_{st}\}}) x_{jst}!}1(x_{jst} > 0)
\]
and expected value
\[
\E_{X|Y,Z}X_{jst} = \frac{\lambda_{st} \exp{\{\lambda_{st} \}}}{\exp{\{ \lambda_{st} \}} - 1}.
\]
% !147
\noindent The joint probability mass function of $X_{jst}, Z_{jst}$ is then 
\begin{equation*}
    f_{X,Z|\boldsymbol{\lambda}}(x_{jst},z_{jst}) = \left\{
    \begin{array}{lr}
      \exp{\{ -\lambda_{st} \}}, & x_{jst}=0 \mbox{ and } z_{jst} = 0 \\
      \frac{\exp{\{-\lambda_{st} \}} \lambda_{st}^{x_{jst}}}{x_{jst}!}, & x_{jst} > 0 \mbox{ and } z_{jst} = 1 \\
      0 & \mbox{otherwise}
    \end{array}
  \right..
\end{equation*}

@@

A more recent method for testing prey preferences was developed by Agusti et al (2003).  This method attempts to test the hypothesis that individuals are eating prey at random using a test based on Monte Carlo simulation (more specificially the bootstrap).  Unfortunately, the method described does not attempt to model the distribution of the number of prey items eaten, as we have done, and therefore we provide a methodological advance on the approach described previously.  The null hypothesis tested by Agusti et all (2003) is that the proportion of prey species $j=1,\ldots,J$ in the mesocosm, denoted $\pi_j$, is equal to the proportion of prey species $j$ in the predator's guts, denoted by $p_j$, at all of the sampling times.  This is tested by repeatedly simulating the gut contents of spiders in each sample under this hypothesis and then comparing summary statistics computed from the observed data and the simulated data to generate a $p$-value.  The problem with this approach is that the relationship between $\pi_j$ and $p_j$ depends on the number of prey items eaten by each predator.  If each predator had eaten exactly one prey item then the null hypothesis of Agusti et al (2003) would be equivalent to random sampling of prey.  However, the relationship between $\pi_j$ and $p_j$ is more complicated when predators can have more than one prey item in their gut.  Given that a predator has $m$ prey items in its gut, randomly sampled from the mesocosm, the marginal probability that at least one item from prey species $j$ is present is $1-(1-\pi_j)^m$.  The marginal probability that an individual has as least one item from prey species $j$ in its gut is then $\sum_{m=1}^{\infty} (1-(1-\pi_j)^m)P(m)$, where $P(m)$ is the probability that a predator has a total of $m$ species in its gut.  The presence of species in a predator's gut contents also cannot be simulated without modeling the number of items in the gut for exactly the same reason.  This demonstrates the importance of knowing the distribution of prey items in the gut in testing the hypothesis of random sampling.  Monte Carlo simulation methods similar to this could be implemented using the output from our model, but these would be computationally intensive in comparison to the likelihood ratio tests which can be computed almost immediately.

@@

The methods we presented offer ecologists a formal statistical framework to model and test predator feeding preferences across an array of time points and prey species. Building from intuitive assumptions, we have developed methodologies for analyzing counts of prey items consumed and have extended these methods for analyzing simple presence/absence data. We have provided computationally efficient methods for fitting these models using the EM algorithm and developed powerful tests for a hierarchy of hypotheses based on the likelihood ratio. We are not aware of previous methods that have offered such abilities. These methods are available to researchers via the R package spiders which is available at \url{http://cran.r-project.org/web/packages/spiders/index.html}.

@@

\noindent  Thus, the asymptotic distribution of $\mathbf{\hat{k}}_{t}$ is $\mathcal{N}(\frac{c_{st}}{c_{\cdot t}}, (\frac{d \mathbf{k}_t}{d \mathbf{c}_t}) \Sigma (\frac{d \mathbf{k}_t}{d \mathbf{c}_t})')$. 
\nocite{Wickham:2011a,Wickham:2011,Bischl:2014,Roualdes:2014}
\bibliographystyle{spbasic}
\bibliography{refs}
\end{document}

>>>>

\documentclass[smallextended]{svjour3}
\input{TexOpts.tex}
\begin{document}
\title{Formal Modelling of Predator Preferences using Molecular Gut-Content Analysis}
\author{Edward A. Roualdes \and Simon J. Bonner \and Thomas D. Whitney \and James D. Harwood}
\institute{E. A. Roualdes 
        \at Department of Statistics, University of Kentucky, Rm.~311 Multidisciplinary Science Building, 725 Rose Street, Lexington, KY 40536-0082, \\\email{edward.roualdes@uky.edu}
        \and S. J. Bonner
        \at Department of Statistical and Actuarial Sciences, University of Western Ontario, 1151 Richmond Street, London, Ontario, N6A 5B7, Canada
        T.D. Whitney 
        \at Department of Entomology, University of Kentucky, Lexington, KY 40546, USA \emph{Present address: Warnell School of Forestry and Natural Resources, University of Georgia, Athens, GA 30602, USA}
        \and
        J.D. Harwood
        \at Department of Entomology, University of Kentucky, Lexington, KY 40546, USA
}

@@

A comprehensive overview by \citet{Lechowicz:1982}, later summarised by \citet{Manly:2002}, details some of the benefits and faults of the most popular indices.  According to these reviews, a majority of the indices give comparable results, save the linear index $L$ by \citet{Strauss:1979}, despite the fact that most of the methods differ by range and linearity of response.  While Lechowicz recommends one index, $E^*$ by \citet{Vanderploeg:1979} as the ``single best'' \citep{Lechowicz:1982}, albeit imperfect, index, \citet{Manly:2002} instead take the approach of excluding the subset of indices which do not ``estimate any biologically meaningful value'' \citep{Manly:2002}.  \citet{Lechowicz:1982} recommends the index $E^*$, an element of the \citet{Manly:2002} suggested indices, because the index value $0$ denotes random feeding, the index has a range restricted to $[-1,1]$ (though $E^*=1$ is nigh impossible), and the index is based on the predator's choice of prey as a function of both the availability of the prey as well as the number of available prey types (assumed known).  The downside to this index is its lack of reasonable statistical properties \citep{Lechowicz:1982}, thus making the computation of standard errors and thus hypothesis testing difficult.  This is, in fact, a common fault amongst most of the indices.

@@

Our model assumes that numbers of each prey species captured and the numbers consumed by an individual predator follow independent Poisson distributions with separate rates that might possibly vary between species and over time.  Further, we are able to estimate the rates of interest even when exact counts of each prey species eaten within any given time period are not observed.  When exact counts are unobserved, we instead rely on the researcher being able to detect prey DNA within the predator's gut \citep{Schmidt:2014,Raso:2014,Madduppa:2014} and make a simple binary conclusion: this predator ate some of that prey species during this time period or did not.

@@

We assume that samples of both the predators and prey are captured from the study area on $T$ occasions.  Depending on the species involved and the design of the study the predators and prey may be sampled in the same way or using different methods.  In the spider experiment described in Section~\ref{sec:data}, for example, prey are captured using pitfall traps that are dispersed throughout the study area whereas the predators (spiders) were captured by hand.  Predators and prey species are collected and counted at each time period.  We denote the number of predators and the number of traps in each time period $t \in \{1, \ldots, T\}$ by $J_t$ and $I_t$, respectively.  Prey species will be indexed by $s \in \{1, \ldots, S \}$.  Let $X_{jst}$ represent the number of prey species $s$ that predator $j$ ate during period $t$, where $j \in \{1, \ldots, J_t\}$.  Let $Y_{ist}$ represent the number of prey species $s$ found in trap $i$ during period $t$, $i \in \{1, \ldots, I_t\}$.

@@

\begin{enumerate}
\item $c_{st} = c$
\item $c_{st} = c_s$
\item $c_{st} = c_t$
\item $c_{st} = c_{st}$
\end{enumerate}

\noindent The first hypothesis states that the relative rates of sampling for the predator and the traps are the same for all species on all occasions.  One imagines this is the case if the prey move randomly and the predator simply eats prey which comes within its reach, thus suggesting no selection for a particular prey item.  The second hypothesis states that predators sample different prey species at different rates, but each rate is steady across time.  This implies that the predator expresses preferences for one prey species over another, but is unresponsive to changes due to time.  Conversely, the third hypothesis implies that each prey species is sampled similarly within each time period, while the rates across time are allowed to change. This would happen if the average amount of prey consumed by the predators changed over time while the relative rates of sampling remained the same, or if there were changes in the efficiency of the traps, or both. The final hypothesis assumes a predator's selection varies by both time and prey species.  This would make sense if environmental and biological variables, such as weather, prey availability, and/or palatability were affecting predators' selection strategies.

@@

Because the five hypotheses are nested, a natural testing order is suggested in Figure~\ref{fig:hier}.  For instance, suppose interest lies in the predator's dietary preferences with respect to prey species $s_1$ and $s_2$ across three time periods.  The hypothesis that the predator has the same preference for the two prey species can be tested with
\begin{align*}
  H_0&: \lambda_t = c_t \gamma_t, \forall t \\  
  H_1&: \lambda_{st} = c_{st}\gamma_{st}.
\end{align*}
The alternative here allows for the predator's preferences to vary independently for the two species and across the three time points. Similarly, the hypotheses
\begin{align*}
  H_0&: \lambda_s = c_s \gamma_s, \forall s \\  
  H_1&: \lambda_{st} = c_{st}\gamma_{st}
\end{align*}
would test if the predator's eating varies by prey species, but not by time, against the most parameter rich model $c_{st}$.  The above two tests should be performed before any of the simpler models are tested.

@@

\noindent Writing all five hypotheses as $\lambda_{st} = c_{st}\gamma_{st}$, we can, in the simplest cases, find analytic solutions for the maximum likelihood estimates (MLEs) of $c_{st}$, $\gamma_{st}$, and by invariance $\lambda_{st}$.  Such solutions exist for the simplest model, $c_{st}=c$, provided that the data are balanced so that $J_t=J$ and $I_t=I$.  The solutions are
\begin{equation*}
  % \hat{\gamma}_{st} = \frac{X_{\cdot st} + Y_{\cdot st}}{J_t + I_t}, \quad 
  % \text{ and } \quad 
  \hat{c} = \frac{I \sum_{s,t} X_{\cdot st}}{J \sum_{s,t} Y_{\cdot st}}
  \quad
  \text{ and }
  \quad
  \hat{\gamma}_{st} = \frac{X_{\cdot st} + Y_{\cdot st}}{I \left( \frac{\sum_{st} X_{\cdot st}}{\sum_{st} Y_{\cdot st}} + 1 \right)},
\end{equation*}
where $X_{\cdot st} = \sum_{j=1}^{J_t}X_{jst}$ and $Y_{\cdot st} = \sum_{i=1}^{I_t} Y_{ist}$.

@@

We denote the binary random variable indicating whether the $j^{th}$ predator did in fact eat at least one individual of prey species $s$ in time period $t$ by $Z_{jst} = 1(X_{jst} > 0)$.  Given the Poisson assumptions above, these variables are independent Bernoulli observations with success probability $p_{st} = P(Z_{jst}=1)= 1-\exp\{-\lambda_{st}\}$.  Despite not observing $X_{jst}$, we can compute maximum likelihood estimates of the parameters $\boldsymbol{\lambda}, \boldsymbol{\gamma}$ through the EM algorithm using the complete data log-likelihood
\[
l_{comp}(\boldsymbol{\lambda}, \boldsymbol{\gamma}) = \log f_{X,Y,Z}(\boldsymbol x, \boldsymbol y, \boldsymbol z|\boldsymbol{\lambda}, \boldsymbol{\gamma}) = \sum_{s,t}^{S,T} \left[ \sum_{j=1}^{J_t} \log f_{X,Z}(x_{jst},z_{jst}|\boldsymbol{\lambda}) + \sum_{i=1}^{I_t}\log f_Y(y_{ist}|\boldsymbol{\gamma}) \right].
\]
% !143, 146
The mass function of $Y_{jst}$ is exactly as in Section~\ref{sec:count} and so we focus on deriving the joint probability mass function of $X_{jst}$ and $Z_{jst}$.  With the distribution of $Z_{jst}$ given above, we can compute $f_{X,Z}(x_{jst},z_{jst}|\boldsymbol{\lambda})$ by noting that $X_{jst}=0$ with probability 1 if $Z_{jst}=0$, and that $[X_{jst}|Z_{jst}=0]$ has a truncated Poisson distribution with mass function
\[
  f_{X|Y,Z,\boldsymbol{\lambda},\boldsymbol{\gamma}}(x_{jst}|z_{jst}) =
  \frac{\exp{\{-\lambda_{st}\}} \lambda_{st}^{x_{jst}}}{(1 - \exp{\{-\lambda_{st}\}}) x_{jst}!}1(x_{jst} > 0)
\]
and expected value
\[
\E_{X|Y,Z}X_{jst} = \frac{\lambda_{st} \exp{\{\lambda_{st} \}}}{\exp{\{ \lambda_{st} \}} - 1}.
\]
% !147
The joint probability mass function of $X_{jst}, Z_{jst}$ is then 
\begin{equation*}
    f_{X,Z|\boldsymbol{\lambda}}(x_{jst},z_{jst}) = \left\{
    \begin{array}{lr}
      \exp{\{ -\lambda_{st} \}}, & x_{jst}=0 \mbox{ and } z_{jst} = 0 \\
      \frac{\exp{\{-\lambda_{st} \}} \lambda_{st}^{x_{jst}}}{x_{jst}!}, & x_{jst} > 0 \mbox{ and } z_{jst} = 1 \\
      0 & \mbox{otherwise}
    \end{array}
  \right..
\end{equation*}

@@

A more recent method for testing prey preferences was developed by \citet{Agusti:2003}.  This method attempts to test the hypothesis that individuals are eating prey at random using a test based on Monte Carlo simulation (more specificially the bootstrap).  Unfortunately, the method described does not attempt to model the distribution of the number of prey items eaten, as we have done, and therefore we provide a methodological advance on the approach described previously.  The null hypothesis tested by \citet{Agusti:2003} is that the proportion of prey species $j=1,\ldots,J$ in the mesocosm, denoted $\pi_j$, is equal to the proportion of prey species $j$ in the predator's guts, denoted by $p_j$, at all of the sampling times.  This is tested by repeatedly simulating the gut contents of spiders in each sample under this hypothesis and then comparing summary statistics computed from the observed data and the simulated data to generate a $p$-value.  The problem with this approach is that the relationship between $\pi_j$ and $p_j$ depends on the number of prey items eaten by each predator.  If each predator had eaten exactly one prey item then the null hypothesis of \citet{Agusti:2003} would be equivalent to random sampling of prey.  However, the relationship between $\pi_j$ and $p_j$ is more complicated when predators can have more than one prey item in their gut.  Given that a predator has $m$ prey items in its gut, randomly sampled from the mesocosm, the marginal probability that at least one item from prey species $j$ is present is $1-(1-\pi_j)^m$.  The marginal probability that an individual has as least one item from prey species $j$ in its gut is then $\sum_{m=1}^{\infty} (1-(1-\pi_j)^m)P(m)$, where $P(m)$ is the probability that a predator has a total of $m$ species in its gut.  The presence of species in a predator's gut contents also cannot be simulated without modeling the number of items in the gut for exactly the same reason.  This demonstrates the importance of knowing the distribution of prey items in the gut in testing the hypothesis of random sampling.  Monte Carlo simulation methods similar to this could be implemented using the output from our model, but these would be computationally intensive in comparison to the likelihood ratio tests which can be computed almost immediately.

@@

The methods we presented offer ecologists a formal statistical framework to model and test predator feeding preferences across an array of time points and prey species. Building from intuitive assumptions, we have developed methodologies for analyzing counts of prey items consumed and have extended these methods for analyzing simple presence/absence data. We have provided computationally efficient methods for fitting these models using the EM algorithm and developed powerful tests for a hierarchy of hypotheses based on the likelihood ratio. We are not aware of previous methods that have offered such abilities. These methods are available to researchers via the R package spiders which is available at \url{http://cran.r-project.org/web/packages/spiders/index.html} \citep{Roualdes:2014}.

@@

\noindent  Thus, the asymptotic distribution of $\mathbf{\hat{k}}_{t}$ is $\mathcal{N}(\frac{c_{st}}{c_{\cdot t}}, (\frac{d \mathbf{k}_t}{d \mathbf{c}_t}) \Sigma (\frac{d \mathbf{k}_t}{d \mathbf{c}_t})')$. 
%\nocite{Wickham:2011a,Wickham:2011,Bischl:2014,Roualdes:2014}
\bibliographystyle{spbasic}
\bibliography{refs}
\end{document}
